Verifying Concurrent Search Structure Templates for C Programs using VST

### 2023-04-05 concurrency/CONCUR 2023/paper.tex
write more lock-coupling and give-up templates

\section{Search Structure Templates}
The C implementation of a node in our template style is:
\begin{lstlisting}[language = C, backgroundcolor=\color{white},   
	commentstyle=\color{codegreen},
	basicstyle=\ttfamily\footnotesize,
	breakatwhitespace=false,  breaklines=true,  keepspaces=true,                                
	showspaces=false,  showstringspaces=false,
	showtabs=false, tabsize=1]
typedef struct tree {int key; void *value; struct tree_t *left, *right;} tree;
typedef struct tree_t {tree *t; lock_t *lock;} tree_t;
typedef struct pn {struct tree_t *p; struct tree_t *n;} pn;
\end{lstlisting}
Each node (of type \lstinline{tree_t}) contains a \lstinline{lock} field that protects the node and a \lstinline{t} field. The \lstinline{t} field can be either \lstinline{NULL}, indicating that the node is a leaf node with no key or value, or it can point to a \lstinline{tree} struct that stores the node's key, value, and child pointers. 
To implement hand-over-hand locking, which guarantees that at least one lock is maintained during traversal between nodes (i.e., the lock for the next node is obtained before releasing the lock for the current node), the \lstinline{pn} structure must contain two nodes of type \lstinline{tree_t}: the \lstinline{p} field, representing the current node, and the \lstinline{n} field, representing the next node.

\subsection{Lock Coupling}
The lock-coupling template employs the hand-over-hand locking scheme to prevent interference from other threads during traversal. Each thread continuously maintains at least one lock during traversal between nodes, ensuring that other threads cannot interfere with the traversal or invalidate the ongoing search. Figure \ref{traverse_lock} presents two versions of the \lstinline{traverse} function: one written in C (Figure \ref{traverse_lock_a}) and the other in an ML-like language (Figure \ref{traverse_lock_b}). The lock-coupling template algorithm, described in an ML-like language, is presented in Figure \ref{traverse_lock_b} \cite{shasha_1988}.  The algorithm operates on the assumption that certain helper function(s) must be provided by the implementation, which satisfy certain criteria.  Such a function, namely \lstinline{findNext}, is utilized by the algorithm. The \lstinline{findNext} function determines the next node $n'$ to be visited based on the current node $n$ and the key $k$ belonging to the set of key values.

\begin{figure}[!ht]
	\begin{subfigure}[t]{0.48\textwidth}
		\lstinputlisting[language=C, style=myStyle]{lock_traverse.c} 
		\caption{The \lstinline{traverse} method of the lock-coupling template give-up template algorithm written in C}
		\label{traverse_lock_a}
	\end{subfigure}\qquad
	\begin{subfigure}[t]{0.45\textwidth}
		\lstinputlisting[language=caml, style=myStyle]{lock_traverse.ml} 
		\caption{The \lstinline{traverse} method of the lock-coupling template algorithm written in ML-like language}
		\label{traverse_lock_b}	
	\end{subfigure}
	\caption{The \lstinline{traverse} method of the lock-coupling template algorithms}
	\label{traverse_lock}
\end{figure}

>>>>

\section{Search Structure Templates}
The C implementation of a node in our template style is:
\begin{lstlisting}[language = C, backgroundcolor=\color{white},   
	commentstyle=\color{codegreen},
	basicstyle=\ttfamily\footnotesize,
	breakatwhitespace=false,  breaklines=true,  keepspaces=true,                                
	showspaces=false,  showstringspaces=false,
	showtabs=false, tabsize=1]
typedef struct tree {int key; void *value; struct tree_t *left, *right;} tree;
typedef struct tree_t {tree *t; lock_t *lock;} tree_t;
typedef struct pn {struct tree_t *p; struct tree_t *n;} pn;
\end{lstlisting}
Each node (of type \lstinline{tree_t}) contains a \lstinline{lock} field that protects the node and a \lstinline{t} field. The \lstinline{t} field can be either \lstinline{NULL}, indicating that the node is a leaf node with no key or value, or it can point to a \lstinline{tree} struct that stores the node's key, value, and child pointers. 
To implement hand-over-hand locking, which guarantees that at least one lock is maintained during traversal between nodes (i.e., the lock for the next node is obtained before releasing the lock for the current node), the \lstinline{pn} struct contain two nodes of type \lstinline{tree_t}: the \lstinline{p} field, representing the current node, and the \lstinline{n} field, representing the next node.

\subsection{Lock Coupling}
The lock-coupling template employs the hand-over-hand locking scheme to prevent interference from other threads during traversal. Each thread continuously maintains at least one lock during traversal between nodes, ensuring that other threads cannot interfere with the traversal or invalidate the ongoing search. Figure \ref{traverse_lock} presents two versions of the \lstinline{traverse} function: one written in C (Figure \ref{traverse_lock_a}) and the other in an ML-like language (Figure \ref{traverse_lock_b}). The lock-coupling template algorithm, described in an ML-like language, is presented in Figure \ref{traverse_lock_b} \cite{shasha_1988}.  The algorithm operates on the assumption that certain helper function(s) must be provided by the implementation, which satisfy certain criteria.  Such a function, namely \lstinline{findNext}, is utilized by the algorithm. The \lstinline{findNext} function determines the next node $n'$ to be visited based on the current node $n$ and the key $k$ belonging to the set of key values. To utilize the template methodology for concurrent data structure implementations in C, we implement the \lstinline{traverse} function in C, as depicted in Figure \ref{traverse_lock_a}. The \lstinline{pn} struct include two fields: \lstinline{p} and \lstinline{n}, which represent the current and next node, respectively. Additionally, we require a target key \lstinline{k}, and a dummy node's value \lstinline{value} (which we call dummy because it is not actually used in \lstinline{traverse}, but we include it to maintain consistent function signatures with other implementations).

\begin{figure}[!ht]
	\begin{subfigure}[t]{0.48\textwidth}
		\lstinputlisting[language=C, style=myStyle]{lock_traverse.c} 
		\caption{The \lstinline{traverse} method of the lock-coupling template algorithm written in C}
		\label{traverse_lock_a}
	\end{subfigure}\qquad
	\begin{subfigure}[t]{0.45\textwidth}
		\lstinputlisting[language=caml, style=myStyle]{lock_traverse.ml} 
		\caption{The \lstinline{traverse} method of the lock-coupling template algorithm written in ML-like language}
		\label{traverse_lock_b}	
	\end{subfigure}
	\caption{The \lstinline{traverse} method of the lock-coupling template algorithms}
	\label{traverse_lock}
\end{figure}
